Can Symmetric Cryptography Be Liberated from the Von Neumann Style?

### 2024-10-13 paper.tex
typo ?
typos

For all intent and purposes, these type constructors shall remain
abstract throughout this article: we interact with them solely through
their algebraic interface. The first of which is the categorical
notion of functor:
%
\coqfrom{Algebra}{Functor}

To the disciples of Reynolds and Girard, this signature has come to
mean ``parametric data container'': how would you go about asserting
that the type constructor \coqe!Rows.T! is not doing something
non-trivial with the type it is provided as an argument? In other
word, how to be sure that a \coqe!Rows.T bool! behaves ``similarly''
as a \coqe!Rows.T nat!? One could try and argue that Coq does not
allow pattern matching on types but it is rather unsavory to involve
the design of the whole programming language into this argument.

@@

Taking the monoid to be \coqe!list A! (under nil and concatenation,
\ie{} the free monoid construction), we can read off the
directionality we were after.
%
More usefully for our purposes, we can derive generic left and right
rotation rotation operators for any foldable functor \coqe!T! Naperian
over a type \coqe!Ix!:
%
\coqfrom{Rows}{ror_sig}
\coqfrom{Rows}{rol_sig}

@@

Consequently, our hands are forced when it comes to lifting up
individual components:
%
\coqfrom{Skinny/Spec}{round}
%
In particular, we have that \coqe!subcells_! is iterated pointwise
across rows and columns, \coqe!add_round_key_! handles the entire cube
as the support for a Boolean algebra, \coqe!shiftrows_! applicatively
applies its column transformation over individual rows and
\coqe!mixcolumns! handles horizontal faces of the cube as Boolean
algebras.

>>>>

For all intents and purposes, these type constructors shall remain
abstract throughout this article: we interact with them solely through
their algebraic interface. The first of which is the categorical
notion of functor:
%
\coqfrom{Algebra}{Functor}

To the disciples of Reynolds and Girard, this signature has come to
mean ``parametric data container'': how would you go about asserting
that the type constructor \coqe!Rows.T! is not doing something
non-trivial with the type it is provided as an argument? In other
words, how to be sure that a \coqe!Rows.T bool! behaves ``similarly''
as a \coqe!Rows.T nat!? One could try and argue that Coq does not
allow pattern matching on types but it is rather unsavory to involve
the design of the whole programming language into this argument.

@@

Taking the monoid to be \coqe!list A! (under nil and concatenation,
\ie{} the free monoid construction), we can read off the
directionality we were after.
%
More usefully for our purposes, we can derive generic left and right
rotation operators for any foldable functor \coqe!T! Naperian
over a type \coqe!Ix!:
%
\coqfrom{Rows}{ror_sig}
\coqfrom{Rows}{rol_sig}

@@

Consequently, our hands are tied when it comes to lifting up
individual components:
%
\coqfrom{Skinny/Spec}{round}
%
In particular, we have that \coqe!subcells_! is iterated pointwise
across rows and columns, \coqe!add_round_key_! handles the entire cube
as the support for a Boolean algebra, \coqe!shiftrows_! applicatively
applies its column transformation over individual rows and
\coqe!mixcolumns! handles horizontal faces of the cube as Boolean
algebras.
